---
title: "LAB 9"
author: "Emily Carreiro"
date: "3/25/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

```{r}
report_03_11_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>%
  rename(Country.Region = "Country/Region", Province.State = "Province/State")
report_03_11_2020
```
```{r}
head(report_03_11_2020)
str(report_03_11_2020)
```
###Confirmed US Cases March 11th
```{r}
report_03_11_2020 %>% 
  filter (Country.Region == "US") %>% 
  ggplot(aes(x = Confirmed,  y = reorder(Province.State, Confirmed))) + 
    geom_point() +
    ggtitle("Confirmed cases for each US State") +
    ylab("Country/Region") +
    xlab("Confirmed Cases")
```
###Summarize the Data for Graphing Places with Multiple Provinces
```{r}
report_03_11_2020 %>% 
  group_by(Country.Region) %>% 
  summarise(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths))
```
###The Graph
```{r}
report_03_11_2020 %>% 
  group_by(Country.Region) %>% 
  summarise(Deaths = sum(Deaths)) %>% 
  arrange(desc(Deaths)) %>% 
  slice(1:20) %>%
  ggplot(aes(x =reorder(Country.Region,Deaths),  y = Deaths)) + 
    geom_bar(stat = 'identity') +
    ggtitle("The 20 countries with the most reported COV19-related deaths") +
    ylab("Country/Region") +
    xlab("Deaths")+
  coord_flip()
```
###Confirmed Cases
```{r}
time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")
head(time_series_confirmed)
```
###Convert to long format
```{r}
time_series_confirmed_long <- time_series_confirmed %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 

head(time_series_confirmed_long)
```
###Adding another set of data
```{r}
time_series_deaths <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

time_series_deaths_long <- time_series_deaths %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Deaths")
head(time_series_deaths_long)
```
###New Column Being Made
```{r}
time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
head(time_series_confirmed_long)

time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)
```
###Unite the Table
```{r}
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key"))
head(time_series_long_joined)
```
###Checking the NA Values
```{r}
which(is.na(time_series_long_joined$Confirmed))
which(is.na(time_series_long_joined$Deaths))
```
###Now For Recovered
```{r}
### download the file
time_series_recovered <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

### convert to long format
time_series_recovered_long <- time_series_recovered %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Recovered")

### Create the Key
time_series_recovered_long <- time_series_recovered_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Recovered)

### Merge with prior table (only this time will delete the Key column
### because we have no further use for it)
time_series_long_joined <- full_join(time_series_long_joined,
              time_series_recovered_long, by = c("Key")) %>% 
    select(-Key)
head(time_series_long_joined)
```
###Looking at the Combined Data
```{r}
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)

time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
head(time_series_long_joined_counts)
```





###Resulting Graph
```{r}
time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Confirmed)) + 
    geom_point() +
    geom_line() +
    ggtitle("US Confirmed COVID-19 Cases")
```
###Some other countries
```{r}
time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
  filter (Country_Region %in% c("China","Japan", "Korea, South",
                                "Italy","Spain", "US")) %>% 
    ggplot(aes(x = Date,  y = Confirmed)) + 
    geom_point() +
    geom_line() +
    ggtitle("Confirmed COVID-19 Cases") +
    facet_wrap(~Country_Region, ncol=2, scales="free_y")

```
###Several Countires
```{r}
time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
    filter (Country_Region %in% c("China","France","Italy", 
                                "Korea, South", "US")) %>% 
    ggplot(aes(x = Date,  y = Confirmed, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("Confirmed COVID-19 Cases")
```
###Showing All
```{r}
time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = log2(Counts), fill = Report_Type, color = Report_Type)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Cases")
```

###Visualizing Map
```{r}
(echo=FALSE)
library(maps)
library(viridis)
world <- map_data("world")
mybreaks <- c(1, 20, 100, 1000, 50000)
```
###Plot of Data
```{r}
world <- map_data("world")
mybreaks <- c(1, 20, 100, 1000, 50000)
ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_point(data=time_series_confirmed, aes(x=Long, y=Lat, size=`2/25/20`, color=`2/25/20`),stroke=F, alpha=0.7) +
scale_size_continuous(name="Cases", trans="log", range=c(1,7),breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
   scale_alpha_continuous(name="Cases", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="Cases", trans="log",breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme(
    legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA)
  )
```
###Adaptation of code from Prof. Chris Sunderland
```{r}
time_series_confirmed_long2 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
    rename(Province.State = "Province/State", Country.Region = "Country/Region") %>%
    pivot_longer(-c(Province.State, Country.Region, Lat, Long),
    names_to = "Date", values_to = "cumulative_cases") %>%
    mutate(Date = mdy(Date) - days(1),
        Place = paste(Lat,Long,sep="_")) %>%
    group_by(Place,Date) %>%
        summarise(cumulative_cases = ifelse(sum(cumulative_cases)>0,
        sum(cumulative_cases),NA_real_),
        Lat = mean(Lat),
        Long = mean(Long)) %>%
    mutate(Pandemic_day = as.numeric(Date - min(Date)))
head(time_series_confirmed_long2)
```
###Plotting the Data
```{r}
static <- ggplot(subset(time_series_confirmed_long2,Date %in% seq(min(Date),max(Date),7)),
            aes(x = Long, y = Lat, size = cumulative_cases/1000)) +
            borders("world", colour = NA, fill = "grey90") +
            theme_bw() +
            geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5) +
            labs(title = 'COVID-19 spread',x = '', y = '',
                 size="Cases (x1000))") +
            theme(legend.position = "right") +
            coord_fixed(ratio=1)+
            facet_wrap(.~Date,nrow=3)
static
```
###Exercises:
###Graphs for Daily Reports
```{r}
report_03_24_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-24-2020.csv")) %>%
rename(Country.Region = "Country_Region", Province.State = "Province_State")
head(report_03_24_2020)
```

###Graph1: Confirmed Cases in US States
```{r}
report_03_24_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-24-2020.csv")) %>%
  rename(Country_Region = "Country_Region", Province_State = "Province_State")
head(report_03_24_2020)
```

```{r}
report_03_24_2020 %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarise(Confirmed=sum(Confirmed))%>% 
  ggplot(aes(x = Confirmed,  y = reorder(Province_State, Confirmed))) + 
    geom_point() +
    ggtitle("Confirmed cases for each US State") +
    ylab("State") +
    xlab("Confirmed Cases") +
    theme(axis.text.y = element_text (size = 5 ))
```
###Graph 2:	Massachusetts Provinces Confirmed Cases
```{r}
report_03_24_2020 %>% 
  filter (Province_State == "Massachusetts") %>% 
  group_by(Confirmed) %>% 
  ggplot(aes(y = Confirmed, x = reorder(Admin2, Confirmed),fill=Admin2, Confirmed)) + 
  coord_flip()+
    geom_bar(stat = 'identity') +
   scale_fill_hue()+
    ggtitle("Confirmed cases for each US State") +
    ylab("Country/Region") +
    xlab("Confirmed Cases")
```
###Exercise 2: Time Series
```{r}
time_series_recovered <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>%
  rename(Province.State = "Province/State", Country.Region = "Country/Region")
head(time_series_recovered)
```
###Converting Format
```{r}
time_series_recovered_long <- time_series_recovered %>% 
               pivot_longer(-c(Province.State, Country.Region, Lat, Long),
                            names_to = "Date", values_to = "Recovered") %>% 
               group_by(Country.Region,Date) %>% 
               summarise(Recovered = sum(Recovered))


time_series_recovered_long$Date <- mdy(time_series_recovered_long$Date)
head(time_series_recovered_long)
```

###Graph 
```{r}
time_series_recovered_long %>% 
  filter (Country.Region == "US") %>% 
    ggplot(aes(x = Date,  y = Recovered, color=Recovered)) + 
    geom_point() +
    geom_line() +
    ggtitle("US Confirmed COVID-19 Recovery Cases")
```
###Graph of Other Countries 
```{r}
time_series_recovered_long %>% 
  filter (Country.Region %in% c("China","Spain","Italy", 
                                "Korea, South", "Iran")) %>% 
    ggplot(aes(x = Date,  y = Recovered, color = Country.Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("Confirmed COVID-19 Recovery Cases")
```
###Exercises 3: Maps
```{r}
world <- map_data("world")
mybreaks <- c(1, 20, 100, 1000, 50000)
ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_point(data=time_series_recovered, aes(x=Long, y=Lat, size=`2/25/20`, color=`2/25/20`),stroke=F, alpha=0.7) +
  scale_size_continuous(name="Cases", trans="log", range=c(1,7),breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  # scale_alpha_continuous(name="Cases", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="Cases", trans="log",breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme(
    legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA)
  )
```
###Map2
```{r}
time_series_recovered_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>%
    rename(Province.State = "Province/State", Country.Region = "Country/Region") %>%
    pivot_longer(-c(Province.State, Country.Region, Lat, Long),
    names_to = "Date", values_to = "cumulative_cases") %>%
    mutate(Date = mdy(Date) - days(1),
        Place = paste(Lat,Long,sep="_")) %>%
    group_by(Place,Date) %>%
        summarise(cumulative_cases = ifelse(sum(cumulative_cases)>0,
        sum(cumulative_cases),NA_real_),
        Lat = mean(Lat),
        Long = mean(Long)) %>%
    mutate(Pandemic_day = as.numeric(Date - min(Date)))
head(time_series_confirmed_long2)
```

```{r}
statica <- ggplot(subset(time_series_recovered_long,Date %in% seq(min(Date),max(Date),7)),
            aes(x = Long, y = Lat, size = cumulative_cases/1000)) +
            borders("world", colour = NA, fill = "grey90") +
            theme_bw() +
            geom_point(shape = 21, color='blue', fill='blue', alpha = 0.5) +
            labs(title = 'COVID-19 Recovery',x = '', y = '',
                 size="Cases (x1000))") +
            theme(legend.position = "right") +
            coord_fixed(ratio=1)+
            facet_wrap(.~Date,nrow=3)
statica
```







